---
title: "Binomial Distribution - R Tutorial"
date: November 18, 2017
author: Jason Seto
output:
  html_notebook:
    theme: united
    toc: yes
    toc_float: yes
  html_document:
    toc: yes
---

Link: http://www.r-tutor.com/elementary-statistics/probability-distributions/binomial-distribution

# Overview

> The binomial distribution is a discrete probability distribution. It describes the outcome of n independent trials in an experiment. Each trial is assumed to have only two outcomes, either success or failure. If the probability of a successful trial is p, then the probability of having x successful outcomes in an experiment of n independent trials is as follows.

* Discrete probability distribution is used to describe binary outcomes based in independent trials

## Problem

> Suppose there are twelve multiple choice questions in an English class quiz. Each question has five possible answers, and only one of them is correct. Find the probability of having four or less correct answers if a student attempts to answer every question at random.

* What is the probability of having 4 or less correct answers on 12 multiple choice questions where there are 5 possible answers for each question?

## Solution

* To find the probability of having EXACTLY 4 correct

dbinom(<Number of binary outcomes>, size = <number of tries>, prob=< probability of success each time>)

```{r}
dbinom(4, size=12, prob=0.2)
```

* To find the probability of having 4 OR LESS orrect, you would add the successive probabilies of fewer successes

```{r}
dbinom(0, size=12, prob=0.2) +
dbinom(1, size=12, prob=0.2) +
dbinom(2, size=12, prob=0.2) +
dbinom(3, size=12, prob=0.2) +
dbinom(4, size=12, prob=0.2)
```

* The likelihood of getting 4 or less correct is 92.7%

You can also leverage the cumulative probability function

pbinom(<probability that you have this number of successes or less>, size=<number of tries>, prob=<likely success at each try> )

```{r}
pbinom(4,size=12,prob=.2)
```
* The likliehood of getting four or less correct is 92.7%